---
title: "Barbell lifts prediction"
author: "Gil Huesca"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, we use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to create a model to predict how well a persone perform barbell lifts.

The training data for this project are available here: [training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data are available here: [test data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

For more information about the dataset and its project, you can consult the document in 

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013

or this [link](http://groupware.les.inf.puc-rio.br/har).

## Data preparation

Training and test sets are read from file in this section. Also they are prepared according to the following elements:
* The sets include values as NA, #DIV/0! and the empty string. Those are transformed in NA values.
* The first 7 columns ae general identification values that are not interesting for the purpose of this project. Those columns are taken out from both sets.
* There are columns that posses NA values, so they have incomplete information. These columns are taken out from both sets.

```{r cars}
trainSet <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
trainSet <- trainSet[,-(1:7)]

testSet <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
testSet <- testSet[,-(1:7)]

aux <- c(which(apply(is.na(trainSet), 2, any)),which(apply(is.na(testSet), 2, any)))

trainSet <- trainSet[,-aux]
testSet <- testSet[,-aux]

```